---
title: "Manual Curation Counts"
author: "Nate Olson"
date: "12/21/2018"
output: "pdf_document"
---

```{r}
library(tidyverse)
```

Loaded cleaned version of table sent by Aaron Wenger.
```{r}
dat <- read_csv("data/manual_curation/justin_table_mod - Sheet1.csv") %>%
    filter(`Var #` != 0) %>% 
    ## homopol col annotation
    mutate(homopol_TR = if_else(str_detect(Notes, "homopol|TR"), "Yes","No"))
```
Total number of variants
```{r}
nrow(dat)
```
Benchmark is correct for 30/31 in homopolymers
DV/CCS is correct for 20/29 outside homopolymers, 14 of which are in LINE
```{r}
dat  %>% 
    group_by(`Line Element`, `Correct\n Call`, homopol_TR) %>% 
    summarise(count = n()) %>% 
    arrange(homopol_TR, `Line Element`)
```
Should be 30 inside homopolymers

Original table used to generate variant callset for manual curation
```{r}
dat2 <- read_csv("data/manual_curation/hg002-ccs-deepvar-curate-JZ - November 19, 2_38 AM - hg002-ccs-deepvar-curate-JZ.csv") %>% 
    ## Removing extra line 
    filter(POS != 232051483	| VAR != "GCA/G") %>% 
    ## Matching variant notation with dat
    mutate(VAR = if_else(POS == 232051483, "GCA/GCATCATGGAGAATGGGACATCTC", VAR))
```

The homopol and target_cat annotations should match unless the methods we used to filter out homopolymers missed some. 

```{r}
dat %>% 
    rename(VAR = "Variant", CHROM = "Chr", POS = "Position") %>% 
    left_join(dat2) %>% 
    group_by(homopol_TR, target_cat) %>% 
    summarise(count = n())
```


Variants no in homopolymers CCS errors due to neighboring homopolymers
3	37083407
6	41984320

Variants in homopolymers not in our target regions, which were suppose to include all homopolymers and tandem repeats
10	29087199 between a 3 bp T homopolymer and 6 bp C homopolymer not in bed files used to identify our target regions
8	132575025 between 2 bp C homopolymer and 5 bp A homopolymer


Variants not identified as in homopolymers or tandem repeats in manual curation notes.
```{r}
dat %>% 
    rename(VAR = "Variant", CHROM = "Chr", POS = "Position") %>% 
    left_join(dat2) %>% 
    group_by(homopol_TR, target_cat) %>% 
    filter(homopol_TR == "No", target_cat == "target")
```
Two variants, one correctly called by PacBio and one correctly called by GIAB. 
__TODO__ want to verify that they are actually in homo or TR


Variants described as near or in homopolymers or tandem repeats in manual curation notes but not in target regions.
```{r}
dat %>% 
    rename(VAR = "Variant", CHROM = "Chr", POS = "Position") %>% 
    left_join(dat2) %>% 
    group_by(homopol_TR, target_cat) %>% 
    filter(homopol_TR == "Yes", target_cat == "non-target")
```
8 variants in this category, 4 called correctly by GIAB and 4 by PacBio.   
The 4 correctly called by GIAB are errors in the PacBio callsets due to neighboring homopolymers, and are all indels. 
The 4 correctly called by PacBio are errors due to long tandem repeats not in the GA4GH TRs. 

```{r}
dat %>% 
    rename(VAR = "Variant", CHROM = "Chr", POS = "Position") %>% 
    left_join(dat2) %>% 
    # colnames()
    group_by(homopol_TR, target_cat, `Correct\n Call`, `Line Element`) %>% 
    summarise(count = n()) %>% 
    arrange(target_cat, `Line Element`, `Correct\n Call`, homopol_TR) %>% 
    spread(`Correct\n Call`, count, fill = 0)
```

```{r}
dat %>% 
    rename(VAR = "Variant", CHROM = "Chr", POS = "Position") %>% 
    left_join(dat2) %>% 
    group_by(homopol_TR, target_cat) %>% 
    filter(`Correct\n Call` == "GIAB", target_cat == "non-target")
```

